Exploring the conformational dynamics of alanine dipeptide in solution subjected to an external electric field: A nonequilibrium molecular dynamics simulation

### 2014-03-03 writedown/npt-neq/alanine.tex
Luigi's revise, send to Giovanni and Christof

\title{Exploring the conformational dynamics of alanine dipeptide in solution subjected to an external electric field: A nonequilibrium molecular dynamics simulation}
\author{Han Wang}
\email{han.wang@fu-berlin.de}
\affiliation{Institute for Mathematics, Freie Universit\"at Berlin, Germany}
\author{Christof Sch\"utte}
\affiliation{Institute for Mathematics, Freie Universit\"at Berlin, Germany}
\affiliation{Zuse Institute Berlin, Germany}
\author{Giovanni Ciccotti}
% \affiliation{Department of Physics, University of Roma "La Sapienza", Italy}
\affiliation{School of Physics, Room 302B UCD-EMSC, University College Dublin, Belfield, Dublin 4, Ireland}
\affiliation{Dipartimento di Fisica, Universit\`a ``La Sapienza'' and CNISM, Piazzale Aldo Moro 5, 00185 Roma, Italy}
\author{Luigi Delle Site}
\affiliation{Institute for Mathematics, Freie Universit\"at Berlin, Germany}

@@

In the present paper, we study a small peptide, alanine dipeptide, as a model for showing how the method can be used. 
The choice of a small peptide, rather than a larger one, is done firstly because it
is easier to draw clear conclusions on the nonequilibrium properties
of a single peptide by excluding the slow-going interplay between
different peptide segments along a chain. 
Secondly, the computational cost required is rather massive and thus  testing
the various methodology proposed here would be hampered for large molecules,
although they remain our goal.  In any case, alanine dipeptide
is large enough so that the conformational changes of its internal degrees of freedom represent valid observables for a test of validity and utility of our approach.
The paper is organized as follows:
  We firstly describe the idea of D-NEMD, and propose a technical improvement consisting in the definition and application of a local thermostating process.
Next, after introducing the physical
observables chosen to analyze the nonequilibrium response,
the numerical results for the case of a static and oscillatory EF, respectively, are discussed. By calculating the nonequilibrium properties from D-NEMD
simulations, we are able to address
a few relevant questions: (i)~How the conformation of the peptide
changes in an EF (conformational space);
(ii)~When these changes happen (time scale),
and how likely they are (probability). 
To our knowledge, this study is one of the few (if not the first)  where
time-dependent nonequilibrium w.r.t. conformation
is explicitly considered for a (relatively large) molecule
in solution subject to an external time-dependent EF.

@@

In this section we propose a procedure to solve the problem: we
start observing that the dynamics without thermostat of the alanine, 
and the water molecules in the closer solvation shells, is
crucial to determine the response of the observable of interest
under the EF perturbation
(in this work, the observable is the conformational change of alanine).
Instead the detailed dynamics of
the water molecules far away is not of primary importance: the region beyond the first solvation shells plays mainly the role of a thermodynamic bath. 
Therefore, instead of simulating an infinitely
large system, we simulate our finite size system with periodic
boundary condition, and with the following characteristics:
(i)~The simulating region is divided in two
subregions, a spherical region around the alanine dipeptide of radius $R_{ex}$,
centered at the alpha-carbon. Here the dynamics is
not subject to a thermostat.
We refer to such subregion as \emph{the  dynamical
region}. (ii)~Beyond $R_{ex}$, the dynamics of the water molecules is coupled
to a Langevin thermostat.  This region is called \emph{the thermostated
region}. For technical convenience we
fix the position of the alpha-carbon in space, so that the alanine is always
located at the center of the simulation region (see Fig.~\ref{fig:tmp2}).
In these conditions, the  dynamics is preserved in the dynamical region (where the properties of interest are observed), while 
possible artificial effects of the thermostating process in the outside region
are negligible due to the finite correlation
range of liquid water.
At the same time, the thermostated region works as a infinitely large
environment that effectively absorbs the extra heat produced by the EF.
The validity of the above
statements will be checked by a series of numerical tests where it is shown that
the response of observables in nonequilibrium (i.e. under the effect of the EF) does not depend on the size of the system
and on the size of the  dynamical subregion, provided that they are
reasonably large.

\recheck{ The proposed temperature control method is very similar to
  the stochastic boundary condition proposed by Brooks~III and Karplus~\cite{brooks1983deformable}.
  The only difference is that we do not explicitly consider a boundary
  region: The system that is divided into a dynamical region (reaction region in their paper) and
  a thermostated regoin (stochastic buffer region in their paper)
  is put into a 3-D cubic periodic simulation box.
}

@@

\subsection{System settings and simulation protocol}
The system is set up in a $2.7\times 2.7\times
2.7\, \textsf{nm}^3$ periodic simulation region, with one alanine dipeptide
described by the CHARMM27 force field~\cite{foloppe2000all}, and dissolved in 644 TIP3P~\cite{jorgensen1983comparison}
water molecules.  \recheck{The grid-based energy correction map (CMAP)~\cite{mackerell2004extending}
for the backbone dihedral angles is also taken into consider.}
%The EF exserts force $\vect F = q \vect E(t)$ on
%the partial charge $q$ in the system.
The size of the  dynamical region is $R_{ex} = 1.0$~nm.
All simulations are performed by a home-modified GROMACS~4.5~\cite{pronk2013gromacs} with support to the CHARMM
force field~\cite{bjelkmar2010implementation}.
First, an equilibrium NVT simulation at 300~K of
100~\textsf{ns} was performed with a Langevin thermostat (time-scale
$\tau_T = 0.5~\textsf{ps}$).  Along the trajectory, configurations were taken every
50~\textsf{ps} and we used 2000 initial configurations for each nonequilibrium
MD simulation (if not stated otherwise for specific cases).
The branching trajectories were integrated by the
Leap-frog scheme with the aforementioned nonequilibrium
temperature control technique.  The
time step was $0.002~\textsf{ps}$.
\recheck{The short-range
interaction (van der Waals interaction)
had the cut-off radius of 1.0~nm, and was smoothed from $0.8$ to $1.0$~nm by the
``switch'' method provided by GROMACS.
A energy conserving particle mesh Ewald method (PME)~\cite{darden1993pme, essmann1995spm}
method was applied to calculate the electrostatic interaction in this
periodic system. The the direct space part of PME was cut-offed and smoothed in the
same way as the van der Waals interaction.}
\recheck {All hydrogen involving bonds are constrained by LINCS~\cite{hess1997lincs}, expect that
  the TIP3P water molecules are constrained by SETTLE algorithm~\cite{miyamoto2004settle}.}
In the thermostated region, the original dynamics was
coupled to a Langevin thermostat with $\tau_T = 0.1~\textsf{ps}$.
  In all testing cases, this local thermostat is able to control
  the system at the desired temperature, i.e.~300~K (the results
  are not presented in this paper).
The whole system is also coupled to a Parrinello-Rahman barostat~\cite{parrinello1981polymorphic} with $\tau_P = 2.0~\textsf{ps}$ to keep
the pressure at ambient condition (1~Bar). Since the
change of the system size is small and slow, the pressure control
does not have an sizable effect on the
dynamics of the system.
At time $t=0$~ps, the system
has been fully equilibrated without any EF. From $t=0$ to
the warm-up time $t=t_{\textrm{warm}}$, the EF is switched on linearly, while
after $t=t_{\textrm{warm}}$, the field is kept constant in time at
$E_{\infty}$.  In this work we consider $t_{\textrm{warm}} = 10$~ps,
and $E_{\infty} = 1$~V/nm.
The direction of the field is arbitrarily chosen along the
$x$ direction.
In this paper we denote the
EF as a function of time $\vect E(t)$.
Therefore, for the case of constant EF, we have
$\vect E(t) = (E_\infty\cdot t/t_{warm},0,0)$ for $0\leq t < t_{warm}$, and 
$\vect E(t) = (E_\infty,0,0)$
for $t \geq t_{warm}$.

@@

From Fig.~\ref{fig:tmp4}, it is evident that the probability with which a
conformation occurs changes when the external EF is applied.
The most evident case is conformation $\confc$; in fact it is {almost} not present before the external field is applied but appears in a clear way as a response to the action of the EF. Following this line of thought, we are interested to study the change of conformation under the action of EF. Specifically, for each conformational change, we will analyze the relation between the structural relaxation of the molecule and its corresponding time scale. 
The starting point is the calculation of the probability
of each conformation:
\begin{align}
  P_I(t) = \mathbb P ( (\phi_t,\psi_t) \in I), \quad  I \in \{\confaa, \confab, \confba, \confbb, \confc\},
\end{align}
Notice that the probability is time dependent, being an observable in nonequilibrium situation.
Next,
we consider the net probability flux from conformation $I$ to $J$, defined by:
\begin{align}\nonumber
  F_{J,I}(t) & = \frac1{\Delta t}\,
  \bigg[\,
  \mathbb P \big( (\phi_{t-\Delta t},\psi_{t-\Delta t}) \in J,\, (\phi_t, \psi_t) \in I \,\big) 
  -
  \mathbb P \big( (\phi_{t-\Delta t}, \psi_{t-\Delta t}) \in I,\, (\phi_t, \psi_t) \in J \,\big)
  \bigg], \\\label{eqn:tmp7}
  & J,I \in \{\confaa, \confab, \confba, \confbb, \confc\}.
\end{align}
% {\bf Here for $\lim_{\Delta \to 0}$ is intended that in the simulation the time interval is taken large enough so that we have  enough events and thus stable numbers (see also Ref.~\cite{schuette2011markov})}.
Where $\mathbb P\big( (\phi_{t-\Delta t}, \psi_{t-\Delta t}) \in J,\, (\phi_t, \psi_t) \in I\,\big)$
is the joint probability of the alanine being in conformation $J$ at
time $(t-\Delta t)$ \emph{and} being in conformation $I$ at time $t$.
The same definition applies to
$\mathbb P \big( (\phi_{t-\Delta t}, \psi_{t-\Delta t}) \in I,\, (\phi_t, \psi_t) \in J\,\big)$.
Here the time interval $\Delta t$ should be small enough compared to the
time scale of conformational dynamics,
so that the changes in the conformation probabilities
can be treated in linear approximation.
At the same time,  $\Delta t$ should be also large enough compared with
the timestep of the MD integrator, 
so that the quantities of Eq.~\eqref{eqn:tmp7} are well-defined, and can be
estimated with sufficient numerical accuracy (see also Ref.~\cite{schuette2011markov}).
\recheck{In all the numerical examples of this work, $\Delta t = 1$~ps is used}.
A positive value of $F_{J,I}(t)$ indicates that the
net flux goes from $J$ to $I$,
while a negative value indicates a net flux from $I$ to $J$.
Through $F_{J,I}(t)$, the analysis of the nonequilibrium process 
is projected onto the analysis of the probabilistic link between discretized conformations $\{\confaa, \confab, \confba, \confbb, \confc\}$.

\recheck{ The definition of the probability flux is similar to the
  kinetic rate.  However, we want to stress that, in general, the
  probability flux should not be interpreted as a kinetic rate, and
  one cannot replace the nonequilibrium molecular dynamics by a
  kinetic analysis. The main reason is, firstly, that the nonequilibrium
  nature of the problem asks for a small time interval $\Delta t$ so that the probability flux is well defined.
  Secondly, within the small
  time interval (we use 1~ps), the conformational
  dynamics of alanine dipeptide is generally not Markovian~\cite{schuette2011markov, prinz2011markov},
  therefore, a direct application of the kinetic
  analysis is questionable.
  % Since the main goal of our research is to propose a
  % nonequilibrium molecular dynamics simulation method, the numerical
  % results will not be compared with the kinetic analysis.
}

In the case of oscillatory EF, the behaviour of the net probability flux is also highly oscillatory, thus, in order to clearly identify the essential features of the molecular conformational changes, it is convenient to study the integrated net probability flux, defined as:
\begin{align}\label{eqn:tmp7a}
  Q_{J,I} (t) = \int_0^t F_{J,I}(\tau)d \tau,
\end{align}
which expresses the cumulative effects due to the action of EF on the conformations of the molecule.

@@

\begin{figure}
  \centering
  \includegraphics[]{fig.ala.new/fig-meta-more.eps}
  \caption{The time-dependent probability of conformations $P_I(t)$ 
    when a constant EF is present, where $\sum_I P_I(t) = 1, \forall t$.
    The warm-up time is $t_{warm} = 10$~ps. The red line refers to conformation $I = \confaa$,
    green to $\confab$, dark blue to $\confba$, purple to $\confbb$ and light blue
    to $\confc$. \recheck{
      The solid lines denote the results of one branching trajectory
      starting from each initial conformation, while the dashed lines denote
      the results of four branching trajectories (different random seeds for the Langevin thermostat) starting from each initial
      conformation.
      The shadow region with each solid line denotes the
      statistical uncertainty of the result at 95\% confidence level.
    }
  }
  \label{fig:tmp5}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=0.3\textwidth]{fig.ala.new/fig-trans-010-flux-1.eps}
  \includegraphics[width=0.3\textwidth]{fig.ala.new/fig-trans-010-flux-2.eps}
  \includegraphics[width=0.3\textwidth]{fig.ala.new/fig-trans-010-flux-3.eps}
  \includegraphics[width=0.3\textwidth]{fig.ala.new/fig-trans-010-flux-4.eps}
  \includegraphics[width=0.3\textwidth]{fig.ala.new/fig-trans-010-flux-5.eps}
  \caption{
    The net probability flux for the constant EF case.
    The warm-up time is $t_{warm} = 10$~ps.
    The probability flux $F_{J,I}(t)$ (defined by Eqn.~\eqref{eqn:tmp7}) is reported in units of $\textrm{ps}^{-1}$.
    From top, left to right we report $I = \confaa$, $\confab$, $\confba$; bottom, left to right $\confbb$ and
    $\confc$, respectively. In each plot, the red line stands for $J=\confaa$,
    green for $J=\confab$, dark blue for $J=\confba$, purple for $J=\confbb$ and light blue
    for $J=\confc$.
    }
  \label{fig:tmp6}
\end{figure}

The time-dependent probability density of the system to stay in a certain
conformation is given in
Fig.~\ref{fig:tmp5}.
\recheck{The simulation results of starting one branching trajectory
  from each initial conformation are compared with those starting
  four branching trajectories from each initial conformation (each with
  different random seed for the Langevin thermostat).
  They are consistent within the statistical uncertainty, which
  means our simulation result is robust with respect to the randomness
  introduced by the (local) Langevin thermostat.
}
The $\beta$-sheet conformations $\confba$ and $\confbb$,
which are present at the initial stage after EF is switched on,
\recheck{fade away as the system relaxes}. The
probability of the $\alpha$-helix conformation $\confaa$, grows from 0.45 to 0.69.
The $\alpha$-helix $\confab$  does not change significantly  under the action of EF.
The left-hand helix conformation $\confc$ noticeably grows from 0 to 0.22.
We plot the net probability fluxes in Fig.~\ref{fig:tmp6}.
Results show that for short $t$, the net fluxes are generally non-zero,
however, 
as $t$ goes to infinity,
all fluxes converge to zero. This indicates
that as the EF is switched on, the system first is driven away from the initial
equilibrium (at zero EF);  This is a nonequilibrium process.
\recheck{After a time long enough, the system
is fully relaxed to the new conformations,
and stays there as long as the EF is kept. However, whether the new conformations
are in the new equilibrium defined the constant EF is not clear, because
there may exist very slow nonequilibrium processes, for example as long as 1~$\mu$s,
which cannot be discovered by our nonequilibrium simulation of only 1~ns.
What we can do by the proposed nonequilibrium simulation
method is to study time-dependent behaviours that are no longer than
the simulation time. 
}
% After a time long enough, the system
% is fully relaxed to the new equilibrium defined by the action of the constant EF,
% and stays there as long as the EF is kept.
An important point is that despite the system is finally relaxed to the new conformation,
the relaxation timescales corresponding to each conformation are rather different. 
From Fig.~\ref{fig:tmp5} and \ref{fig:tmp6}, we observe mainly
three different timescales 10~ps, 100~ps and 500~ps,
whose corresponding conformations are summarized in Tab.~\ref{tab:tmp1}.
This is a relevant result because it shows the possibility of employing an external EF as a tool to identify relevant time scales in the conformational behaviour of a molecule in solution.

@@

\subsection{Molecular dipole response}
The fact that our external perturbation corresponds to the action of
an electric field, naturally leads to the question of how
\recheck{
  the molecular dipole aligns along the direction of external EF}, and, in turn, how this is related
to the overall \recheck{conformational selection} of the molecule reported in
the previous section.
\begin{figure}
  \centering
  \includegraphics[width=0.45\textwidth]{fig.ala.new/field.dipole/fig-dipol-order-conf.eps}
  \caption{The
    $x$-component of the molecular dipole moment, and the orientational order parameter
    of the alanine molecule as a
    function of time. The solid line represents the dipole moment,
    while the dashed line represents the orientational order parameter.
    The left vertical axis is for the dipole moment, while
    the right is for the orientational order parameter.
  }
  \label{fig:tmp3}
\end{figure}

@@

\begin{figure}
  \centering
  \includegraphics[]{fig.ala.new/fig-meta-conv.eps}
  \caption{Testing the finite-size effect on the dynamical and thermostated region.
    Probabilities of different conformations are presented versus time.
    See the text for more details.
    \recheck{The shadow region denotes the
      statistical uncertainty of the $L=2.7$~nm $R_{ex}=1.0$~nm case at 95\% confidence level.
    }
  }
  \label{fig:tmp7}
\end{figure}

@@

% \begin{thebibliography}{10}

% \bibitem{bohr2000microwave}
% Henrik Bohr and Jakob Bohr.
% \newblock Microwave-enhanced folding and denaturation of globular proteins.
% \newblock {\em Physical Review E}, 61(4):4310, 2000.

% \bibitem{bohr2000microwave-1}
% Henrik Bohr and Jakob Bohr.
% \newblock Microwave enhanced kinetics observed in ord studies of a protein.
% \newblock {\em Bioelectromagnetics}, 21(1):68--72, 2000.

% \bibitem{dePomerai2000cell}
% David de~Pomerai, Clare Daniells, Helen David, Joanna Allan, Ian Duce, Mohammed
%   Mutwakil, David Thomas, Phillip Sewell, John Tattersall, Don Jones, and Peter
%   Candido.
% \newblock Cell biology: Non-thermal heat-shock response to microwaves.
% \newblock {\em Nature}, 405:417--418, 2000.

% \bibitem{inskip2001cellular}
% Peter~D Inskip, Robert~E Tarone, Elizabeth~E Hatch, Timothy~C Wilcosky,
%   William~R Shapiro, Robert~G Selker, Howard~A Fine, Peter~M Black, Jay~S
%   Loeffler, and Martha~S Linet.
% \newblock Cellular-telephone use and brain tumors.
% \newblock {\em New England Journal of Medicine}, 344(2):79--86, 2001.

>>>>

\title{Exploring the conformational dynamics of alanine dipeptide in solution subjected to an external electric field: A nonequilibrium molecular dynamics simulation}
\author{Han Wang}
\email{han.wang@fu-berlin.de}
\affiliation{Institute for Mathematics, Freie Universit\"at Berlin, Germany}
\author{Christof Sch\"utte}
\affiliation{Institute for Mathematics, Freie Universit\"at Berlin, Germany}
\affiliation{Zuse Institute Berlin, Germany}
\author{Giovanni Ciccotti}
% \affiliation{Department of Physics, University of Roma "La Sapienza", Italy}
\affiliation{School of Physics, University College Dublin, Belfield, Dublin 4, Ireland}
\affiliation{Dipartimento di Fisica, Universit\`a ``La Sapienza'' and CNISM, Piazzale Aldo Moro 5, 00185 Roma, Italy}
\author{Luigi Delle Site}
\affiliation{Institute for Mathematics, Freie Universit\"at Berlin, Germany}

@@

In the present paper, we study a small peptide, alanine dipeptide, as a model for showing how the method can be used. 
The choice of a small peptide, rather than a larger one, is done firstly because it
is easier to draw clear conclusions on the nonequilibrium properties
of a single peptide by excluding the slow-going interplay between
different peptide segments along a chain. 
Secondly, the computational cost required is rather massive and thus  testing
the various methodology proposed here would be hampered for large molecules,
although they remain our goal.  In any case, alanine dipeptide
is large enough so that the conformational changes of its internal degrees of freedom represent valid observables for a test of validity and utility of our approach.
The paper is organized as follows:
  We firstly describe the idea of D-NEMD, and propose a technical improvement consisting in the definition and application of a local thermostating process.
Next, after introducing the physical
observables chosen to analyze the nonequilibrium response,
the numerical results for the case of a static and oscillatory EF, respectively, are discussed. By calculating the nonequilibrium properties from D-NEMD
simulations, we are able to address
a few relevant questions: (i)~How the conformation of the peptide
changes in an EF (conformational space);
(ii)~When these changes happen (time scale),
and how likely they are (probability).
\recheck{One may wonder whether the information extracted from the D-NEMD simulation could have been
extracted just as well from conventional equilibrium simulations.
  In the appendix we show that in this case the D-NEMD approach shall be preferred to the equilibrium approach. The argument is provided by the comparison between the results of the nonequilibrium method and
  those from equilibrium simulations, within the linear response theory, based on the analysis of correlation functions.}
To our knowledge, this study is one of the few (if not the first)  where
time-dependent nonequilibrium w.r.t. conformation
is explicitly considered for a (relatively large) molecule
in solution subject to an external time-dependent EF.

@@

In this section we propose a procedure to solve the problem: we
start observing that the dynamics without thermostat of the alanine, 
and the water molecules in the closer solvation shells, is
crucial to determine the response of the observable of interest
under the EF perturbation
(in this work, the observable is the conformational change of alanine).
Instead the detailed dynamics of
the water molecules far away is not of primary importance: the region beyond the first solvation shells plays mainly the role of a thermodynamic bath. 
Therefore, instead of simulating an infinitely
large system, we simulate our finite size system with periodic
boundary condition, and with the following characteristics:
(i)~The simulating region is divided in two
subregions, a spherical region around the alanine dipeptide of radius $R_{ex}$,
centered at the alpha-carbon. Here the dynamics is
not subject to a thermostat.
We refer to such subregion as \emph{the  dynamical
region}. (ii)~Beyond $R_{ex}$, the dynamics of the water molecules is coupled
to a Langevin thermostat.  This region is called \emph{the thermostated
region}. For technical convenience we
fix the position of the alpha-carbon in space, so that the alanine is always
located at the center of the simulation region (see Fig.~\ref{fig:tmp2}).
In these conditions, the  dynamics is preserved in the dynamical region (where the properties of interest are observed), while 
possible artificial effects of the thermostating process in the outside region
are negligible due to the finite correlation
range of liquid water.
At the same time, the thermostated region works as a infinitely large
environment that effectively absorbs the extra heat produced by the EF.
The validity of the above
statements will be checked by a series of numerical tests where it is shown that
the response of observables in nonequilibrium (i.e. under the effect of the EF) does not depend on the size of the system
and on the size of the  dynamical subregion, provided that they are
reasonably large.
\recheck{ The proposed temperature control method has similarities with 
  the stochastic boundary condition proposed by Brooks and Karplus~\cite{brooks1983deformable}.
  However, differently from their approach, we do not explicitly consider a boundary region: The system is instead divided into a dynamical region (corresponding to the reaction region of \cite{brooks1983deformable}) and
  a thermostated region (corresponding to the stochastic buffer region in \cite{brooks1983deformable}) in a 3-D cubic periodic simulation box.
}

@@

\subsection{System settings and simulation protocol}
The system is set up in a $2.7\times 2.7\times
2.7\, \textsf{nm}^3$ periodic simulation region, with one alanine dipeptide
described by the CHARMM27 force field~\cite{foloppe2000all}, and dissolved in 644 TIP3P~\cite{jorgensen1983comparison}
water molecules.  \recheck{The grid-based energy correction map (CMAP)~\cite{mackerell2004extending}
for the backbone dihedral angles is also used.}
%The EF exserts force $\vect F = q \vect E(t)$ on
%the partial charge $q$ in the system.
The size of the  dynamical region is $R_{ex} = 1.0$~nm.
All simulations are performed by a home-modified GROMACS~4.5~\cite{pronk2013gromacs} with support to the CHARMM
force field~\cite{bjelkmar2010implementation}.
First, an equilibrium NVT simulation at 300~K of
100~\textsf{ns} was performed with a Langevin thermostat (time-scale
$\tau_T = 0.5~\textsf{ps}$).  Along the trajectory, configurations were taken every
50~\textsf{ps} and we used 2000 initial configurations for each nonequilibrium
MD simulation (if not stated otherwise for specific cases).
The branching trajectories were integrated by the
Leap-frog scheme with the aforementioned nonequilibrium
temperature control technique.  The
time step was $0.002~\textsf{ps}$.
\recheck{The short-range
interaction (van der Waals interaction)
have a cut-off radius of 1.0~nm, and has been smoothed from $0.8$ to $1.0$~nm by the
``switch'' method provided by the GROMACS code.
A energy conserving particle mesh Ewald method (PME)~\cite{darden1993pme, essmann1995spm}
method was applied to calculate the electrostatic interaction in this
periodic system. For the direct space part of PME the cut-off and smoothing follow the same principles as those applied to the van der Waals interaction.}
\recheck {All hydrogen involving bonds are constrained by LINCS~\cite{hess1997lincs}, except 
  the TIP3P water molecules which are constrained by the SETTLE algorithm~\cite{miyamoto2004settle}.}
In the thermostated region, the original dynamics was
coupled to a Langevin thermostat with $\tau_T = 0.1~\textsf{ps}$.
  In all testing cases, this local thermostat is able to control
  the system at the desired temperature, i.e.~300~K (the results
  are not presented in this paper).
The whole system is also coupled to a Parrinello-Rahman barostat~\cite{parrinello1981polymorphic} with $\tau_P = 2.0~\textsf{ps}$ to keep
the pressure at ambient condition (1~Bar). Since the
change of the system size is small and slow, the pressure control
does not have an sizable effect on the
dynamics of the system.
At time $t=0$~ps, the system
has been fully equilibrated without any EF. From $t=0$ to
the warm-up time $t=t_{\textrm{warm}}$, the EF is switched on linearly, while
after $t=t_{\textrm{warm}}$, the field is kept constant in time at
$E_{\infty}$.  In this work we consider $t_{\textrm{warm}} = 10$~ps,
and $E_{\infty} = 1$~V/nm.
The direction of the field is arbitrarily chosen along the
$x$ direction.
In this paper we denote the
EF as a function of time $\vect E(t)$.
Therefore, for the case of constant EF, we have
$\vect E(t) = (E_\infty\cdot t/t_{warm},0,0)$ for $0\leq t < t_{warm}$, and 
$\vect E(t) = (E_\infty,0,0)$
for $t \geq t_{warm}$.

@@

From Fig.~\ref{fig:tmp4}, it is evident that the probability with which a
conformation occurs changes when the external EF is applied.
The most evident case is conformation $\confc$; in fact it is {almost} not present before the external field is applied but appears in a clear way as a response to the action of the EF. Following this line of thought, we are interested to study the change of conformation under the action of EF. Specifically, for each conformational change, we will analyze the relation between the structural relaxation of the molecule and its corresponding time scale. 
The starting point is the calculation of the probability
of each conformation:
\begin{align}
  P_I(t) = \mathbb P ( (\phi_t,\psi_t) \in I), \quad  I \in \{\confaa, \confab, \confba, \confbb, \confc\},
\end{align}
Notice that the probability is time dependent, being an observable in nonequilibrium situation.
Next,
we consider the net probability flux from conformation $I$ to $J$, defined by:
\begin{align}\nonumber
  F_{J,I}(t) & = \frac1{\Delta t}\,
  \bigg[\,
  \mathbb P \big( (\phi_{t-\Delta t},\psi_{t-\Delta t}) \in J,\, (\phi_t, \psi_t) \in I \,\big) 
  -
  \mathbb P \big( (\phi_{t-\Delta t}, \psi_{t-\Delta t}) \in I,\, (\phi_t, \psi_t) \in J \,\big)
  \bigg], \\\label{eqn:tmp7}
  & J,I \in \{\confaa, \confab, \confba, \confbb, \confc\}.
\end{align}
% {\bf Here for $\lim_{\Delta \to 0}$ is intended that in the simulation the time interval is taken large enough so that we have  enough events and thus stable numbers (see also Ref.~\cite{schuette2011markov})}.
Where $\mathbb P\big( (\phi_{t-\Delta t}, \psi_{t-\Delta t}) \in J,\, (\phi_t, \psi_t) \in I\,\big)$
is the joint probability of the alanine being in conformation $J$ at
time $(t-\Delta t)$ \emph{and} being in conformation $I$ at time $t$.
The same definition applies to
$\mathbb P \big( (\phi_{t-\Delta t}, \psi_{t-\Delta t}) \in I,\, (\phi_t, \psi_t) \in J\,\big)$.
Here the time interval $\Delta t$ should be small enough compared to the
time scale of conformational dynamics,
so that the changes in the conformation probabilities
can be treated in linear approximation.
At the same time,  $\Delta t$ should be also large enough compared with
the time step of the MD integrator, 
so that the quantities of Eq.~\eqref{eqn:tmp7} are well-defined, and can be
estimated with sufficient numerical accuracy (see also Ref.~\cite{schuette2011markov}).
\recheck{In all the numerical examples of this work, $\Delta t = 1$~ps is used}.
A positive value of $F_{J,I}(t)$ indicates that the
net flux goes from $J$ to $I$,
while a negative value indicates a net flux from $I$ to $J$.
Through $F_{J,I}(t)$, the analysis of the nonequilibrium process 
is projected onto the analysis of the probabilistic link between discretized conformations $\{\confaa, \confab, \confba, \confbb, \confc\}$.
\recheck{ The concept of net probability flux employed by us is similar to the
  concept of kinetic rate.  However, it must be underlined that, in general, the
  probability flux should not be interpreted as a kinetic rate.
  Nonequilibrium molecular dynamics cannot be replaced by 
  kinetic analysis, the main reason is, firstly, that the nonequilibrium
  nature of the problem requires the use of a small time interval $\Delta t$ so that the probability flux is well defined.
  Secondly, within the small
  time interval (we use 1~ps), the conformational
  dynamics of alanine dipeptide is generally not Markovian~\cite{schuette2011markov, prinz2011markov},
  therefore, a direct application of the kinetic
  analysis may be questionable.}
In the case of oscillatory EF, the behaviour of the net probability flux is also highly oscillatory, thus, in order to clearly identify the essential features of the molecular conformational changes, it is convenient to study the integrated net probability flux, defined as:
\begin{align}\label{eqn:tmp7a}
  Q_{J,I} (t) = \int_0^t F_{J,I}(\tau)d \tau,
\end{align}
which expresses the cumulative effects due to the action of EF on the conformations of the molecule.

@@

\begin{figure}
  \centering
  \includegraphics[]{fig.ala.new/fig-meta-more.eps}
  \caption{The time-dependent probability of conformations $P_I(t)$ 
    when a constant EF is present, where $\sum_I P_I(t) = 1, \forall t$.
    The warm-up time is $t_{warm} = 10$~ps. The red line refers to conformation $I = \confaa$,
    green to $\confab$, dark blue to $\confba$, purple to $\confbb$ and light blue
    to $\confc$. \recheck{
      The solid lines denote the results of one branching trajectory
      starting from each initial conformation, while the dashed lines denote
      the results of four branching trajectories (different random seeds for the Langevin thermostat) starting from each initial
      conformation.
      The shadowed region with each solid line denotes the
      statistical uncertainty of the result at 95\% confidence level.
    }
  }
  \label{fig:tmp5}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=0.3\textwidth]{fig.ala.new/fig-trans-010-flux-1.eps}
  \includegraphics[width=0.3\textwidth]{fig.ala.new/fig-trans-010-flux-2.eps}
  \includegraphics[width=0.3\textwidth]{fig.ala.new/fig-trans-010-flux-3.eps}
  \includegraphics[width=0.3\textwidth]{fig.ala.new/fig-trans-010-flux-4.eps}
  \includegraphics[width=0.3\textwidth]{fig.ala.new/fig-trans-010-flux-5.eps}
  \caption{
    The net probability flux for the constant EF case.
    The warm-up time is $t_{warm} = 10$~ps.
    The probability flux $F_{J,I}(t)$ (defined by Eqn.~\eqref{eqn:tmp7}) is reported in units of $\textrm{ps}^{-1}$.
    From top, left to right we report $I = \confaa$, $\confab$, $\confba$; bottom, left to right $\confbb$ and
    $\confc$, respectively. In each plot, the red line stands for $J=\confaa$,
    green for $J=\confab$, dark blue for $J=\confba$, purple for $J=\confbb$ and light blue
    for $J=\confc$.
    }
  \label{fig:tmp6}
\end{figure}
The time-dependent probability density of the system to stay in a certain
conformation is given in
Fig.~\ref{fig:tmp5}.
\recheck{In order to show that  our simulation results are robust with respect to the randomness introduced by the (local) Langevin thermostat, we have tested the effects of changing random seeds for the thermostat in the thermostated region. We compared simulation results obtained by starting a branching trajectory
  from each initial conformation with those obtained from four different branching trajectories from each initial conformation (each of four with
  different random seed for the Langevin thermostat in the thermostated region).
  Results are consistent within the statistical uncertainty (see Fig.\ref{fig:tmp5}).
}
The $\beta$-sheet conformations $\confba$ and $\confbb$,
which are present at the initial stage after EF is switched on,
\recheck{fade away as the system relaxes}. The
probability of the $\alpha$-helix conformation $\confaa$, grows from 0.45 to 0.69.
The $\alpha$-helix $\confab$  does not change significantly  under the action of EF.
The left-hand helix conformation $\confc$ noticeably grows from 0 to 0.22.
We plot the net probability fluxes in Fig.~\ref{fig:tmp6}.
Results show that for short $t$, the net fluxes are generally non-zero,
however, 
as $t$ goes to infinity,
all fluxes converge to zero. This indicates
that as the EF is switched on, the system first is driven away from the initial
equilibrium (at zero EF);  This is a nonequilibrium process.
\recheck{After sufficiently long time, the system
is fully relaxed to the new conformations,
and remains in such a state as long as the EF is switched on. However, on the very long time scale,
there may exist very slow nonequilibrium processes, for example within an 1~$\mu$s range,
which cannot be captured by the duration of our nonequilibrium simulation (1~ns).
Obviously, the proposed nonequilibrium simulation
method can study time-dependent behaviours that are no longer than
the simulation time. 
}
% After a time long enough, the system
% is fully relaxed to the new equilibrium defined by the action of the constant EF,
% and stays there as long as the EF is kept.
An important point is that despite the system is finally relaxed to the new conformation,
the relaxation timescales corresponding to each conformation are rather different. 
From Fig.~\ref{fig:tmp5} and \ref{fig:tmp6}, we observe mainly
three different timescales 10~ps, 100~ps and 500~ps,
whose corresponding conformations are summarized in Tab.~\ref{tab:tmp1}.
This is a relevant result because it shows the possibility of employing an external EF as a tool to identify relevant time scales in the conformational behaviour of a molecule in solution.

@@

\subsection{Molecular dipole response}
The fact that our external perturbation corresponds to the action of
an electric field, naturally leads to the question 
\recheck{
  about the alignment of the molecular dipole vector along the direction of external EF}, and, in turn, how this is related
to the overall \recheck{conformational selection} of the molecule reported in
the previous section.
\begin{figure}
  \centering
  \includegraphics[width=0.45\textwidth]{fig.ala.new/field.dipole/fig-dipol-order-conf.eps}
  \caption{The
    $x$-component of the molecular dipole moment, and the orientational order parameter
    of the alanine molecule as a
    function of time. The solid line represents the dipole moment,
    while the dashed line represents the orientational order parameter.
    The left vertical axis is for the dipole moment, while
    the right is for the orientational order parameter.
  }
  \label{fig:tmp3}
\end{figure}

@@

\begin{figure}
  \centering
  \includegraphics[]{fig.ala.new/fig-meta-conv.eps}
  \caption{Testing the finite-size effect on the dynamical and thermostated region.
    Probabilities of different conformations are presented versus time.
    See the text for more details.
    \recheck{The shadowed region denotes the
      statistical uncertainty of the $L=2.7$~nm $R_{ex}=1.0$~nm case at 95\% confidence level.
    }
  }
  \label{fig:tmp7}
\end{figure}

@@
